---
title: "st499"
author: "Grace Salo"
date: "1/28/2022"
output: html_document
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
```

## Import Data
```{r import}
data <- read.csv("simulated_data.csv", nrows = 60)
colnames(data) <- c('Time', 'Base', 'Noise1', 'Noise5', 'Noise10', 'ReturnBase', 'Return1', 'Return5', 'Return10')
head(data)
tail(data)
    # NOTE: removed last obs to end up with 60 total
```

## EDA
```{r eda}
# Scatterplot: Base, Noise1, Noise5, Noise10 vs. Time
ggplot(data=data) + geom_point(aes(x=Time, y=Noise10), color="purple") + geom_point(aes(x=Time, y=Noise5), color="blue") + geom_point(aes(x=Time, y=Noise1), color="green") + geom_point(aes(x=Time, y=Base), color="yellow") + labs(title="Base, Noise1, Noise5, Noise10 vs. Time", caption="Base=yellow, Noise1=green, Noise5=blue, Noise10=purple")

# Scatterplot: ReturnBase, Return1, Return5, Return10 vs. Time
ggplot(data=data) + geom_point(aes(x=Time, y=Return10), color="purple") + geom_point(aes(x=Time, y=Return5), color="blue") + geom_point(aes(x=Time, y=Return1), color="green") + geom_point(aes(x=Time, y=ReturnBase), color="yellow") + labs(title="ReturnBase, Return1, Return5, Return10 vs. Time", caption="ReturnBase=yellow, Return1=green, Return5=blue, Return10=purple")

# Return5 vs. Time with ReturnBase line of best fit
ggplot(data=data) + geom_point(aes(x=Time, y=Return5)) + geom_line(aes(x=Time, y=ReturnBase), color="blue") + labs(title="Return5 vs. Time with ReturnBase line (blue)")


# print some!
png(file="Base-Noise_vs_Time.png", width=450, height=450)
ggplot(data=data) + geom_point(aes(x=Time, y=Noise10), color="purple") + geom_point(aes(x=Time, y=Noise5), color="blue") + geom_point(aes(x=Time, y=Noise1), color="green") + geom_point(aes(x=Time, y=Base), color="yellow") + labs(title="Base, Noise1, Noise5, Noise10 vs. Time", caption="Base=yellow, Noise1=green, Noise5=blue, Noise10=purple")
dev.off()
```




## Model: ReturnBase ~ Time and Prediction
#### returnbase (not fn) final
```{r returnbase}
# set starting number of obs (n)
n <- 30

# subset data into (1:n) and (n+1, end)
train <- data %>% filter(Time <= n) %>% select(Time, ReturnBase)
test <- data %>% filter(Time > n) %>% select(Time, ReturnBase)

# model ReturnBase ~ Time
mod <- lm(ReturnBase ~ Time, data=train)
summary(mod)
# plot of predicted return values from model
ggplot(data=train) + geom_point(aes(x=Time, y=mod$coefficients[1] + 
  mod$coefficients[2]*Time)) + labs(x=paste0("1st n=",n), y="Predicted values of Return \n from Linear Model", title=paste0("n=",n," predicted return from model"), subtitle="model: ReturnBase ~ Time")

# predict 2nd part of values
pred <- predict(object=mod, newdata=test)

# add column to final dataset with predicted values
final <- data %>% select(Time, Base, ReturnBase) %>% 
  mutate(PredReturnBase = append(data[1:n, 6], pred))

# plot new data with predicted values of ReturnBase in the second half
ggplot(data=final) + geom_point(aes(x=Time, y=PredReturnBase)) + 
  labs(title=paste0("n=",n," original obs + n=",(nrow(data)-n)," predicted obs of Return"), subtitle="model: ReturnBase ~ Time")

# create PredBase from exp(PredReturnBase[i])*Base[i-1]
final <- final %>% mutate(PredBase = exp(PredReturnBase)*lag(Base))
# set first PredBase to original Base value
final$PredBase[1] = final$Base[1]

# plot PredBase vs. Time
ggplot(data=final) + 
  geom_point(aes(x=Time, y=PredBase), color="blue") +
  geom_point(aes(x=Time, y=Base), color="green") + 
  labs(title="Base vs. Time", subtitle="Original (green) and Predicted (blue) data")

# calculate residuals (should be 0 for base)
final <- final %>% mutate(res = PredReturnBase - ReturnBase)

# plot histogram of residuals
ggplot(data=final) + geom_histogram(aes(x=res), bins=nrow(final))

# calculate MAPE



```


#### EXTRA: returnbase function to specify n
```{r returnbase-fn}
baseReturnFn <- function(n) {
  # subset data into (1:n) and (n+1, end)
  train <- data %>% filter(Time <= n) %>% select(Time, ReturnBase)
  test <- data %>% filter(Time > n) %>% select(Time, ReturnBase)
  
  # model ReturnBase ~ Time
  mod <- lm(ReturnBase ~ Time, data=train)
  summary(mod)
  # plot of predicted return values from model
  ggplot(data=train) + geom_point(aes(x=Time, y=mod$coefficients[1] + mod$coefficients[2]*Time)) + labs(x=c("1st n=",n), y="Predicted values of Return \n from Linear Model", title=c("n=",n,"predicted return from model"), subtitle="model: ReturnBase ~ Time")
  
  # predict 2nd part of values
  pred <- predict(object=mod, newdata=test)

  # add column to final dataset with predicted values
  final <- data %>% select(Time, ReturnBase)
  final[3] <- append(data[1:n, 6], pred)
  colnames(final)[3] <- "PredReturnBase"
  
  # calculate residuals (should be 0 for base)
  final <- final %>% mutate(res = PredReturnBase - ReturnBase)
  
  return(final)
}

# call function: baseReturnFn (provide n)
n = 30
baseFinal <- baseReturnFn(n)

# plot new data with predicted values of ReturnBase in the second half
ggplot(data=baseFinal) + geom_point(aes(x=Time, y=PredReturnBase)) + labs(title=paste0("n=",n," original obs + n=",(nrow(data)-n)," predicted obs of Return"), subtitle="model: ReturnBase ~ Time")

# plot histogram of residuals
ggplot(data=baseFinal) + geom_histogram(aes(x=res), bins=nrow(baseFinal))
```

#### EXTRA: observations: n = 30
```{r returnbase-n30, eval=FALSE}
# set starting number of obs (n)
n <- 30

# subset data into (1:n) and (n+1, end)
base.d30_1 <- data %>% filter(Time <= n) %>% select(Time, ReturnBase)
base.d30_2 <- data %>% filter(Time > n) %>% select(Time, ReturnBase)

# model ReturnBase ~ Time
mod <- lm(ReturnBase ~ Time, data=base.d30_1)
summary(mod)
# plot of predicted return values from model
ggplot(data=base.d30_1) + geom_point(aes(x=Time, y=mod$coefficients[1] + mod$coefficients[2]*Time)) + labs(x=c("1st n",n), y="Predicted values of Return \n from Linear Model", title="n=30 predicted return from model", subtitle="model: ReturnBase ~ Time")

# predict 2nd part of values
pred <- predict(object=mod, newdata=base.d30_2)

# add column to final dataset with predicted values
base.d30 <- data %>% select(Time, ReturnBase)
base.d30[3] <- append(data[1:n, 6], pred)
colnames(base.d30)[3] <- "PredReturnBase"
# plot new data with predicted values of ReturnBase in the second half
ggplot(data=base.d30) + geom_point(aes(x=Time, y=PredReturnBase)) + labs(title="n=30 original obs + n=30 predicted obs of Return", subtitle="model: ReturnBase ~ Time")


# double-check that pred and data it's replacing are lined up correctly (yes!)
cbind(data[31:60, 6], pred)

# calculate residuals (should be 0 for base)
base.d30 <- base.d30 %>% mutate(res = PredReturnBase - ReturnBase)
# plot histogram of residuals
ggplot(data=base.d30) + geom_histogram(aes(x=res), bins=nrow(base.d30))
```



## Model: Return1 ~ Time and Prediction
#### EXTRA: return1 function to specify n
```{r return1-fn}
return1Fn <- function(n) {
  # subset data into (1:n) and (n+1, end)
  train <- data %>% filter(Time <= n) %>% select(Time, Return1)
  test <- data %>% filter(Time > n) %>% select(Time, Return1)
  
  # model Return1 ~ Time
  mod <- lm(Return1 ~ Time, data=train)
  summary(mod)
  # plot of predicted return values from model
  ggplot(data=train) + geom_point(aes(x=Time, y=mod$coefficients[1] + mod$coefficients[2]*Time)) + labs(x=c("1st n=",n), y="Predicted values of Return \n from Linear Model", title=c("n=",n,"predicted return from model"), subtitle="model: Return1 ~ Time")
  
  # predict 2nd part of values
  pred <- predict(object=mod, newdata=test)

  # add column to final dataset with predicted values
  final <- data %>% select(Time, Return1)
  final[3] <- append(data[1:n, 6], pred)
  colnames(final)[3] <- "PredReturn1"
  
  # calculate residuals
  final <- final %>% mutate(res = PredReturn1 - Return1)
  
  return(final)
}

# call function: return1Fn (provide n)
n = 15
r1Final <- return1Fn(n)

# plot new data with predicted values of Return1 in the second half
ggplot(data=r1Final) + geom_point(aes(x=Time, y=PredReturn1)) + labs(title=paste0("n=",n," original obs + n=",(nrow(data)-n)," predicted obs of Return"), subtitle="model: Return1 ~ Time")

# plot histogram of residuals
ggplot(data=r1Final) + geom_histogram(aes(x=res), bins=nrow(r1Final))
```


## Model: Return5 ~ Time and Prediction



## Model: Return10 ~ Time and Prediction
#### return10 (not fn) final
```{r return10}
# set starting number of obs (n)
n = 5

# subset data into (1:n) and (n+1, end)
train <- data %>% filter(Time <= n) %>% select(Time, Return10)
test <- data %>% filter(Time > n) %>% select(Time, Return10)

# model Return10 ~ Time
mod <- lm(Return10 ~ Time, data=train)
summary(mod)
# plot of predicted return values from model
ggplot(data=train) + geom_point(aes(x=Time, y=mod$coefficients[1] + 
  mod$coefficients[2]*Time)) + labs(x=paste0("1st n=",n), y="Predicted values of Return \n from Linear Model", title=paste0("n=",n," predicted return from model"), subtitle="model: Return10 ~ Time")


# predict 2nd part of values
pred <- predict(object=mod, newdata=test)

# add column to final dataset with predicted values
final <- data %>% select(Time, Noise10, Return10) %>% 
  mutate(PredReturn10 = append(data[1:n, 6], pred))

# plot new data with predicted values of ReturnBase in the second half
ggplot(data=final) + geom_point(aes(x=Time, y=PredReturn10)) + 
  labs(title=paste0("n=",n," original obs + n=",(nrow(data)-n)," predicted obs of Return"), subtitle="model: Return10 ~ Time")

# create Pred10 from exp(PredReturn10[i])*Noise10[i-1]
final <- final %>% mutate(Pred10 = exp(PredReturn10)*lag(Noise10))
# set first Pred10 to original Noise10 value
final$Pred10[1] = final$Noise10[1]

# plot Pred10 vs. Time
ggplot(data=final) + 
  geom_point(aes(x=Time, y=Pred10), color="blue") +
  geom_point(aes(x=Time, y=Noise10), color="green") + 
  labs(title="Noise10 vs. Time", subtitle="Original (green) and Predicted (blue) data")


# calculate residuals
final <- final %>% mutate(res = PredReturn10 - Return10)

# plot histogram of residuals
ggplot(data=final) + geom_histogram(aes(x=res), bins=nrow(final))
```

#### EXTRA: return10 function to specify n
```{r return10-fn}
return10Fn <- function(n) {
  # subset data into (1:n) and (n+1, end)
  train <- data %>% filter(Time <= n) %>% select(Time, Return10)
  test <- data %>% filter(Time > n) %>% select(Time, Return10)
  
  # model Return10 ~ Time
  mod <- lm(Return10 ~ Time, data=train)
  summary(mod)
  # plot of predicted return values from model
  ggplot(data=train) + geom_point(aes(x=Time, y=mod$coefficients[1] + mod$coefficients[2]*Time)) + labs(x=c("1st n=",n), y="Predicted values of Return \n from Linear Model", title=c("n=",n,"predicted return from model"), subtitle="model: Return10 ~ Time")
  
  # predict 2nd part of values
  pred <- predict(object=mod, newdata=test)

  # add column to final dataset with predicted values
  final <- data %>% select(Time, Return10)
  final[3] <- append(data[1:n, 6], pred)
  colnames(final)[3] <- "PredReturn10"
  
  # calculate residuals
  final <- final %>% mutate(res = PredReturn10 - Return10)
  
  return(final)
}

# call function: return10Fn (provide n)
n=5
r10Final <- return10Fn(n)

# plot new data with predicted values of Return10 in the second half
ggplot(data=r10Final) + geom_point(aes(x=Time, y=PredReturn10)) + labs(title=paste0("n=",n," original obs + n=",(nrow(data)-n)," predicted obs of Return"), subtitle="model: Return10 ~ Time")

# plot histogram of residuals
ggplot(data=r10Final) + geom_histogram(aes(x=res), bins=nrow(r10Final))
```









# EXTRA BELOW
#### EXTRA: observations: n = 25
```{r returnbase-n25, eval=FALSE}
# set starting number of obs (n)
n <- 25

# subset data into (1:n) and (n+1, end)
base.d25_1 <- data %>% filter(Time <= n) %>% select(Time, ReturnBase)
base.d25_2 <- data %>% filter(Time > n) %>% select(Time, ReturnBase)

# model ReturnBase ~ Time
mod <- lm(ReturnBase ~ Time, data=base.d25_1)
summary(mod)
# plot of predicted return values from model
ggplot(data=base.d25_1) + geom_point(aes(x=Time, y=mod$coefficients[1] + mod$coefficients[2]*Time)) + labs(x="1st n=25", y="Predicted values of Return \n from Linear Model", title="n=25 predicted return from model", subtitle="model: ReturnBase ~ Time")

# predict 2nd part of values
pred <- predict(object=mod, newdata=base.d25_2)

# add column to final dataset with predicted values
base.d25 <- data %>% select(Time, ReturnBase)
base.d25[3] <- append(data[1:n, 6], pred)
colnames(base.d25)[3] <- "PredReturnBase"
# plot new data with predicted values of ReturnBase in the second half
ggplot(data=base.d25) + geom_point(aes(x=Time, y=PredReturnBase)) + labs(title="n=25 original obs + n=35 predicted obs of Return", subtitle="model: ReturnBase ~ Time")


# double-check that pred and data it's replacing are lined up correctly (yes!)
cbind(data[26:60, 6], pred)

# calculate residuals (should be 0 for base)
base.d25 <- base.d25 %>% mutate(res = PredReturnBase - ReturnBase)
# plot histogram of residuals
ggplot(data=base.d25) + geom_histogram(aes(x=res), bins=nrow(base.d25))
```
#### EXTRA: observations: n = 20
```{r returnbase-n20, eval=FALSE}
# set starting number of obs (n)
n <- 20

# subset data into (1:n) and (n+1, end)
base.d20_1 <- data %>% filter(Time <= n) %>% select(Time, ReturnBase)
base.d20_2 <- data %>% filter(Time > n) %>% select(Time, ReturnBase)

# model ReturnBase ~ Time
mod <- lm(ReturnBase ~ Time, data=base.d20_1)
summary(mod)
# plot of predicted return values from model
ggplot(data=base.d20_1) + geom_point(aes(x=Time, y=mod$coefficients[1] + mod$coefficients[2]*Time)) + labs(x="1st n=20", y="Predicted values of Return \n from Linear Model", title="n=20 predicted return from model", subtitle="model: ReturnBase ~ Time")

# predict 2nd part of values
pred <- predict(object=mod, newdata=base.d20_2)

# add column to final dataset with predicted values
base.d20 <- data %>% select(Time, ReturnBase)
base.d20[3] <- append(data[1:n, 6], pred)
colnames(base.d20)[3] <- "PredReturnBase"
# plot new data with predicted values of ReturnBase in the second half
ggplot(data=base.d20) + geom_point(aes(x=Time, y=PredReturnBase)) + labs(title="n=20 original obs + n=40 predicted obs of Return", subtitle="model: ReturnBase ~ Time")


# double-check that pred and data it's replacing are lined up correctly (yes!)
cbind(data[21:60, 6], pred)

# calculate residuals (should be 0 for base)
base.d20 <- base.d20 %>% mutate(res = PredReturnBase - ReturnBase)
# plot histogram of residuals
ggplot(data=base.d20) + geom_histogram(aes(x=res), bins=nrow(base.d20))

# calculate MAPE
#...
```

#### EXTRA: observations: n = 30
```{r return1-n30}
# set starting number of obs (n)
n <- 30

# subset data into (1:n) and (n+1, end)
r1.d30_1 <- data %>% filter(Time <= n) %>% select(Time, Return1)
r1.d30_2 <- data %>% filter(Time > n) %>% select(Time, Return1)

# model ReturnBase ~ Time
mod <- lm(Return1 ~ Time, data=r1.d30_1)
summary(mod)
# plot of predicted return values from model
ggplot(data=r1.d30_1) + geom_point(aes(x=Time, y=mod$coefficients[1] + mod$coefficients[2]*Time)) + labs(x="1st n=30", y="Predicted values of Return1 \n from Linear Model", title="n=30 predicted return from model", subtitle="model: PredReturn1 ~ Time")

# predict 2nd half of values
pred <- predict(object=mod, newdata=r1.d30_2)

# add column to d30 with predicted values for last n=30
r1.d30 <- data %>% select(Time, Return1)
r1.d30[3] <- append(data[1:n, 6], pred)
colnames(r1.d30)[3] <- "PredReturn1"
# plot new data with predicted values of Return1 in the second half
ggplot(data=r1.d30) + geom_point(aes(x=Time, y=PredReturn1)) + labs(title="n=30 original obs + n=30 predicted obs of Return", subtitle="model: PredReturn1 ~ Time")


# double-check that pred and data it's replacing are lined up correctly (yes!)
cbind(data[31:60, 6], pred)

# calculate residuals (should be 0)
r1.d30 <- r1.d30 %>% mutate(res = PredReturn1 - Return1)
# plot histogram of residuals
ggplot(data=r1.d30) + geom_histogram(aes(x=res), bins=nrow(r1.d30))
```




# TRASH BELOW

## practice: turn orig --> return and return --> orig
```{r delta}
a <- data %>% filter(Time < 5) %>% select(Base, ReturnBase)

# formula: orig --> return
log(a[2,1]/a[1,1])
log(a[2,1]) - log(a[1,1])
a[2,2] # answer

# formula: return --> orig
exp(a[2,2])*a[1,1]
a[2,1] # answer
```

## attempt: set returnData, then use quote() and bquote() to reference
```{r try}
# set starting number of obs (n)
n = 30
# set returnData
returnData = quote(Return10)  # use with bquote(.(returnData))

# subset data into (1:n) and (n+1, end)
train <- data %>% filter(Time <= n) %>% select(Time, bquote(.(returnData)))
test <- data %>% filter(Time > n) %>% select(Time, bquote(.(returnData)))

# PROBLEM: doesn't work for lm()
```
## attempt: reconstruct adoption curve using fn etc
```{r trash}
# create fn for PredBase
        # PROBLEM: can't pass column/dataframe names as arguments
reconstFn <- function(data, return, orig) {
  a <- list()
  for(i in 1:nrow(data)) {
    if(i==1) {
      b <- data %>% slice(i) %>% mutate(PredBase = {{orig}}) %>% select(PredBase)
      a <- append(a,b)
    }
    else {
      b <- data %>% slice(i) %>% mutate(PredBase = exp({{col}})*lag({{orig}})) %>% select(PredBase) 
      a <- append(a, b)
    }
  }
  data <- mutate(newCol = a)
  return(data)
}
a <- reconstFn(base.d20, PredReturnBase, Base)

reconstFn(base.d20, PredReturnBase, Base)
b <- base.d20  %>% select(PredReturnBase) %>% slice(3) %>% mutate(newCol = exp(PredReturnBase))
base.d20 <- base.d20 %>% mutate(PredBase = reconstFn(data, PredReturnBase))
```

## attempt: Function to specify any starting set (i.e. ReturnBase, Return1) and any n (i.e. 30,25,20)
```{r return-fn}
returnFn1 <- function(data, n, returnData) {
  # subset data into (1:n) and (n+1, end)
  train <- data %>% filter(Time <= n) %>% select(Time, {{returnData}})
  test <- data %>% filter(Time > n) %>% select(Time, {{returnData}})
  returnFn2(data, n, returnData, train, test)
  return()
}
returnFn2 <- function(train, test) {
  
}
a <- returnFn1(data, n=30, ReturnBase)

returnFn <- function(data, n, returnData) {
  print(data[returnData])
}
a <- returnFn(data, n=30, "ReturnBase")
```

#### OPTION: loop observations: n = 30, 25, 20, 15, 10

## attempt: replace original return with pred
```{r 1}
# replace ReturnBase 31:60 with predicted values
d30 <- data %>% select(Time, ReturnBase)
d30[31:60, 2] <- pred1
# plot new data with predicted values of ReturnBase in the second half
ggplot(data=d30) + geom_point(aes(x=Time, y=ReturnBase)) + labs(title="n=30 original obs + n=30 predicted obs of Return", subtitle="model: ReturnBase ~ Time")
```

## attempt: lm(returnbase ~ base)
```{r hey}
# return base vs. original base
mod1 <- lm(ReturnBase ~ Base, data=d30_1)
summary(mod1)
pred1 <- predict(object=mod1, data=d30_2)
pred1 <- append(NA, pred1)

# combine columns: 2nd half of base data + predicted vals
a <- cbind(d30_2$Base, pred1)
colnames(a) <- c("Base", "ReturnBase")

a <- rbind(d30_1, cbind(d30_2$Base, pred1))

a <- cbind(d30_2$Base, pred1)
rbind(d30_1, a, names=c(""))
```

## attempt: lm(returnbase ~ base)
```{r another}
# plan: take data df (with Time) and just add the predicted values into that

# set starting number of obs (n)
n <- 30

# subset data into (1:n) and (n+1, end)
d30_1 <- data %>% filter(Time <= (n+1)) %>% select(Time, Base, ReturnBase)
d30_2 <- data %>% filter(Time > n) %>% select(Time, Base, ReturnBase)
   # NOTE: d30_1 has 1:30 and d30_2 has 30:60

# original dataset
d30_orig <- data %>% select(Time, Base, ReturnBase)

# create model of ReturnBase ~ Base
mod1 <- lm(ReturnBase ~ Base, data=d30_1)
summary(mod1)
# plot of predicted return values from base model
ggplot(data=d30_1) + geom_point(aes(x=Base, y=mod1$coefficients[1] + mod1$coefficients[2]*Base)) + labs(x="1st half of Base data", y="Predicted values of Return Base\n from Linear Model", title="n=30 predicted return from base model")


# predict 2nd half of values
pred1 <- predict(object=mod1, newdata=d30_2)
plot(x=data$Time, y=data$ReturnBase)
plot(x=d30_2$Time, y=pred1)

# replace ReturnBase 31:60 with predicted values
d30_orig[31:60, 3] <- pred1

ggplot(data=d30_orig) + geom_point(aes(x=Time, y=ReturnBase))
```

## attempt: select n=30
```{r 1.1}
# set starting number of obs (n)
n <- 30

# subset data into (1:n) and (n+1, end)
d30_1 <- data[1:n, ]
   # OR: data %>% filter(Time <= n)
d30_2 <- data[-c(1:n), ]
   # OR: data %>% filter(Time > n)

# from first section of data, create slr
#lm(Base ~ Time, data=d30_1)
```

## attempt: select n=30
```{r 1.2}
# set starting number of obs (n)
n <- 30

# subset data into (1:n) and (n+1, end)
d30_1 <- data %>% filter(Time <= n) %>% select(Base, ReturnBase)
d30_2 <- data %>% filter(Time > n) %>% select(Base, ReturnBase)
   # include Time?
```
